### 2019-08-22 poster/ligo.tex

\textbf{Control Configuration}

\begin{minipage}[t]{0.49\linewidth}
  \begin{tikzfigure}[Specifications and weighting functions magnitude used for $\mathcal{H}_\infty$ synthesis]
    \label{fig:ligo_weights}
    \centering
    \includegraphics[scale=1.8]{figs/ligo_weights.pdf}
  \end{tikzfigure}
\end{minipage}\hfill
\begin{minipage}[t]{0.49\linewidth}
  \begin{tikzfigure}[Comparison of the FIR filters (solid) designed in \cite{hua05_low_ligo} with the filters obtained with $\mathcal{H}_\infty$ synthesis (dashed)]
    \label{fig:comp_fir_ligo_hinf}
    \centering
    \includegraphics[scale=1.8]{figs/comp_fir_ligo_hinf.pdf}
  \end{tikzfigure}
\end{minipage}