### 2019-08-22 poster/poster.tex

\documentclass[%
    a0paper,%
    20pt,%
    portait,%
    margin=0mm,%
    innermargin=10mm,%
    blockverticalspace=10mm,%
    colspace=10mm,%
    subcolspace=5mm%
]{tikzposter}

\input{config.tex}

\addbibresource{ref.bib}

\author{Dehaeze Thomas, Verma Mohit and Collette Christophe}
\date{2019-11-06}
\title{Complementary Filters Shaping Using \scalebox{2}{$\mathcal{H}_\infty$} Synthesis}
\institute{}

\useblockstyle{customstyle}

\begin{document}

\maketitle[%
    roundedcorners=2,
    linewidth=2mm,
    innersep=10mm,
    titletotopverticalspace=10mm,
    titletoblockverticalspace=10mm]

\block[]{}{\input{abstract.tex}}

\begin{columns}
\column{0.5}
\block[]{Introduction}{\input{introduction.tex}}
\block[]{Complementary Filters Requirements}{\input{requirements.tex}}

\column{0.5}
\block[]{Complementary Filters Shaping using $\mathcal{H}_\infty$ Synthesis}{\input{hinfinity.tex}}
\block[]{Application: Design of Complementary Filters used in the Active Vibration Isolation System at the LIGO}{\input{ligo.tex}}
\block[]{Conclusion}{\input{conclusion.tex}}
\end{columns}

\block[]{}{\input{reference.tex}}

\end{document}