### 2019-08-22 poster/requirements.tex

\begin{tikzfigure}[Sensor fusion architecture]
  \label{fig:fusion_super_sensor}
  \centering
  \includegraphics[scale=1.8]{figs/fusion_super_sensor.pdf}
\end{tikzfigure}

\begin{tikzfigure}[Sensor fusion architecture with sensor dynamics uncertainty]
  \label{fig:sensor_fusion_dynamic_uncertainty}
  \centering
  \includegraphics[scale=1.8]{figs/sensor_fusion_dynamic_uncertainty.pdf}
\end{tikzfigure}

\begin{tikzfigure}[Uncertainty set of the super sensor dynamics]
  \label{fig:uncertainty_set_super_sensor}
  \centering
  \includegraphics[scale=1.8]{figs/uncertainty_set_super_sensor.pdf}
\end{tikzfigure}