### 2019-09-26 tikz/figs/comp_fir_ligo_hinf.tex
Review - figures are modified

\setlength\fwidth{6.5cm}
\setlength\fheight{7cm}

\begin{tikzpicture}
  \begin{axis}[%
    width=1.0\fwidth,
    height=0.60\fheight,
    at={(0.0\fwidth, 0.35\fheight)},
    scale only axis,
    xmode=log,
    xmin=0.001,
    xmax=1,
    xtick={0.001,0.01,0.1,1},
    xticklabels={{}},
    xminorticks=true,
    ymode=log,
    ymin=0.002,
    ymax=5,
    ytick={0.001, 0.01, 0.1, 1, 10},
    yminorticks=true,
    ylabel={Magnitude},
    xminorgrids,
    yminorgrids,
    legend style={at={(1,0)}, outer sep=2pt, anchor=south east, legend cell align=left, align=left, draw=black, nodes={scale=0.7, transform shape}}
    ]
    \addplot [color=mycolor1, line width=1.5pt]
      table [x=freqs, y=Hhm, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_hinf.csv};
    \addlegendentry{$H_H(s)$ - $\mathcal{H}_\infty$}
    \addplot [color=mycolor1, dashed, line width=1.5pt]
      table [x=freqs, y=Hhm, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_fir.csv};
    \addlegendentry{$H_H(s)$ - FIR}
    \addplot [color=mycolor2, line width=1.5pt]
      table [x=freqs, y=Hlm, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_hinf.csv};
    \addlegendentry{$H_L(s)$ - $\mathcal{H}_\infty$}
    \addplot [color=mycolor2, dashed, line width=1.5pt]
      table [x=freqs, y=Hlm, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_fir.csv};
    \addlegendentry{$H_L(s)$ - FIR}
  \end{axis}

  \begin{axis}[%
    width=1.0\fwidth,
    height=0.3\fheight,
    at={(0.0\fwidth, 0.0\fheight)},
    scale only axis,
    xmode=log,
    xmin=0.001,
    xmax=1,
    xtick={0.001,  0.01,   0.1,     1},
    xminorticks=true,
    xlabel={Frequency [Hz]},
    ymin=-180,
    ymax=180,
    ytick={-180,  -90,    0,   90,  180},
    ylabel={Phase [deg]},
    xminorgrids,
    ]
    \addplot [color=mycolor1, line width=1.5pt, forget plot]
      table [x=freqs, y=Hhp, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_hinf.csv};
    \addplot [color=mycolor1, dashed, line width=1.5pt, forget plot]
      table [x=freqs, y=Hhp, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_fir.csv};
    \addplot [color=mycolor2, line width=1.5pt, forget plot]
      table [x=freqs, y=Hlp, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_hinf.csv};
    \addplot [color=mycolor2, dashed, line width=1.5pt, forget plot]
      table [x=freqs, y=Hlp, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/comp_ligo_fir.csv};
  \end{axis}
\end{tikzpicture}