### 2019-09-26 tikz/figs/hinf_synthesis_results.tex

\setlength\fwidth{6.5cm}
\setlength\fheight{6cm}

\begin{tikzpicture}
  \begin{axis}[%
    width=1.0\fwidth,
    height=0.5\fheight,
    at={(0.0\fwidth, 0.5\fheight)},
    scale only axis,
    xmode=log,
    xmin=0.1,
    xmax=1000,
    xtick={0.1, 1, 10, 100, 1000},
    xticklabels={{}},
    xminorticks=true,
    ymode=log,
    ymin=0.0005,
    ymax=20,
    ytick={0.001, 0.01, 0.1, 1, 10},
    yminorticks=true,
    ylabel={Magnitude},
    xminorgrids,
    yminorgrids,
    ]
    \addplot [color=mycolor1, line width=1.5pt, forget plot]
    table [x=freqs, y=H1, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/hinf_filters_results.csv};

    \addplot [color=mycolor2, line width=1.5pt, forget plot]
    table [x=freqs, y=H2, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/hinf_filters_results.csv};

    \addplot [color=mycolor1, dashed, line width=1.5pt, forget plot]
    table [x=freqs, y=W1, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/hinf_weights.csv};

    \addplot [color=mycolor2, dashed, line width=1.5pt, forget plot]
    table [x=freqs, y=W2, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/hinf_weights.csv};
  \end{axis}

  \begin{axis}[%
    width=1.0\fwidth,
    height=0.45\fheight,
    at={(0.0\fwidth, 0.0\fheight)},
    scale only axis,
    xmode=log,
    xmin=0.1,
    xmax=1000,
    xtick={0.1, 1, 10, 100, 1000},
    xminorticks=true,
    xlabel={Frequency [Hz]},
    ymin=-200,
    ymax=200,
    ytick={-180,  -90,    0,   90,  180},
    ylabel={Phase [deg]},
    xminorgrids,
    legend style={at={(1,1.1)}, outer sep=2pt , anchor=north east, legend cell align=left, align=left, draw=black, nodes={scale=0.7, transform shape}},
    ]
    \addlegendimage{color=mycolor1, dashed, line width=1.5pt}
    \addlegendentry{$W_1^{-1}$};
    \addlegendimage{color=mycolor2, dashed, line width=1.5pt}
    \addlegendentry{$W_2^{-1}$};
    \addplot [color=mycolor1, line width=1.5pt]
    table [x=freqs, y=H1p, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/hinf_filters_results.csv};
    \addlegendentry{$H_1$};
    \addplot [color=mycolor2, line width=1.5pt]
    table [x=freqs, y=H2p, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/hinf_filters_results.csv};
    \addlegendentry{$H_2$};
  \end{axis}
\end{tikzpicture}