### 2019-09-26 tikz/figs/ligo_weights.tex

\setlength\fwidth{6.5cm}
\setlength\fheight{4cm}

\begin{tikzpicture}
  \begin{axis}[%
    width=1.0\fwidth,
    height=1.0\fheight,
    at={(0.0\fwidth, 0.0\fheight)},
    scale only axis,
    separate axis lines,
    every outer x axis line/.append style={black},
    every x tick label/.append style={font=\color{black}},
    every x tick/.append style={black},
    xmode=log,
    xmin=0.001,
    xmax=1,
    xminorticks=true,
    xlabel={Frequency [Hz]},
    every outer y axis line/.append style={black},
    every y tick label/.append style={font=\color{black}},
    every y tick/.append style={black},
    ymode=log,
    ymin=0.005,
    ymax=20,
    yminorticks=true,
    ylabel={Magnitude},
    axis background/.style={fill=white},
    xmajorgrids,
    xminorgrids,
    ymajorgrids,
    yminorgrids,
    legend style={at={(0,1)}, outer sep=2pt, anchor=north west, legend cell align=left, align=left, draw=black, nodes={scale=0.7, transform shape}}
    ]

    \addplot [color=mycolor1, line width=1.5pt]
      table [x=freqs, y=wHm, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/ligo_weights.csv};
    \addlegendentry{$|w_H|^{-1}$}

    \addplot [color=mycolor2, line width=1.5pt]
      table [x=freqs, y=wLm, col sep=comma] {/home/thomas/Cloud/These/Papers/dehaeze19_desig_compl_filte/matlab/mat/ligo_weights.csv};
    \addlegendentry{$|w_L|^{-1}$}

    \addplot [color=black, dotted, line width=1.5pt]
    table[row sep=crcr]{%
      0.0005	0.008\\
      0.008   0.008\\
    };
    \addlegendentry{Specifications}

    \addplot [color=black, dotted, line width=1.5pt, forget plot]
    table[row sep=crcr]{%
      0.008	0.008\\
      0.04	1\\
    };
    \addplot [color=black, dotted, line width=1.5pt, forget plot]
    table[row sep=crcr]{%
      0.04  3\\
      0.1   3\\
    };
    \addplot [color=black, dotted, line width=1.5pt]
    table[row sep=crcr]{%
      0.1	0.045\\
      2   0.045\\
    };
  \end{axis}
\end{tikzpicture}