### 2019-10-23 tikz/figs/weight_formula.tex

\setlength\fwidth{6.5cm}
\setlength\fheight{3.5cm}

\begin{tikzpicture}
  \begin{axis}[%
    width=1.0\fwidth,
    height=1.0\fheight,
    at={(0.0\fwidth, 0.0\fheight)},
    scale only axis,
    xmode=log,
    xmin=0.1,
    xmax=100,
    xtick={0.1,1,10, 100},
    xminorticks=true,
    ymode=log,
    ymin=0.0005,
    ymax=20,
    ytick={0.001, 0.01, 0.1, 1, 10},
    yminorticks=true,
    ylabel={Magnitude},
    xlabel={Frequency [Hz]},
    xminorgrids,
    yminorgrids,
    ]

    \addplot [color=black, line width=1.5pt, forget plot]
    table [x=freqs, y=ampl, col sep=comma] {/home/thomas/Cloud/thesis/papers/dehaeze19_desig_compl_filte/matlab/matweight_formula.csv};

    \addplot [color=black, dashed, line width=1.5pt]
    table[row sep=crcr]{%
      1     10\\
      100   10\\
    };
    \addplot [color=black, dashed, line width=1.5pt]
    table[row sep=crcr]{%
      0.1  0.001\\
      3    0.001\\
    };

    \addplot [color=black, line width=1.5pt]
    table[row sep=crcr]{%
      0.1  1\\
      100  1\\
    };

    \addplot [color=black, dashed, line width=1.5pt]
    table[row sep=crcr]{%
      10  2\\
      10  1\\
    };

    \node[below] at (2, 10) {$G_\infty$};
    \node[above] at (2, 0.001) {$G_0$};

    \node[branch] at (10, 2){};
    \draw[dashed, line cap=round] (7, 2) -- (20, 2) node[right]{$G_c$};
    \draw[dashed, line cap=round] (10, 2) -- (10, 1) node[below]{$\omega_c$};

    \node[right] at (3, 0.1) {$+n$};

  \end{axis}
\end{tikzpicture}