### 2019-10-30 poster/conclusion.tex
Poster: Worked on the ligo part

%%% Local Variables:
%%% TeX-master: "poster"
%%% End:

Complementary filters can be used to \textbf{combine multiple sensors} in order
to obtain a \textbf{super sensor}.
Specification on the \textbf{super sensor noise} and on the \textbf{robustness
  of the sensor fusion} are linked to the \textbf{norm of the complementary filters}.
A synthesis method that permits the \textbf{shaping of the complementary filters
  norms} has been proposed and has been successfully applied for the design of
complex filters.