### 2019-10-30 poster/poster.tex
Poster: Add corresponding Author

\documentclass[%
    a1paper,%
    17pt,%
    portait,%
    margin=0mm,%
    innermargin=5mm,%
    blockverticalspace=5mm,%
    colspace=5mm,%
    subcolspace=2mm%
]{tikzposter}

\usetitlestyle[]{Wave}

\input{config.tex}

\addbibresource{ref.bib}

\author[1,2,$\dagger$]{Dehaeze Thomas}
\author[1,3]{Verma Mohit}
\author[1,3]{Collette Christophe}
\affil[1]{Precision Mechatronics Laboratory, A\&M Department, University of Liege, Belgium}
\affil[2]{European Synchrotron Radiation Facility, Grenoble, France}
\affil[3]{BEAMS Department, Free University of Brussels, Belgium}
\affil[$\dagger$]{Corresponding Author. Email: {\tt thomas.dehaeze@esrf.fr}\vspace{-2em}}

\date{2019-11-06}
\title{Complementary Filters Shaping Using $\mathcal{H}_\infty$ Synthesis}
\institute{}

\useblockstyle{customstyle}

\begin{document}

\maketitle[%
    roundedcorners=2,
    linewidth=2mm,
    innersep=5mm,
    titletotopverticalspace=10mm,
    titletoblockverticalspace=10mm]

\block[]{}{\input{abstract.tex}}

\begin{columns}
\column{0.5}
\block[]{Introduction}{\input{introduction.tex}}
\block[]{Sensor Fusion Architecture}{\input{architecture.tex}}
\block[]{Complementary Filters Requirements}{\input{requirements.tex}}

\column{0.5}
\block[]{Complementary Filters Shaping using $\mathcal{H}_\infty$ Synthesis}{\input{hinfinity.tex}}
\block[]{Design of Complementary Filters used in the Active Vibration Isolation System at the LIGO}{\input{ligo.tex}}
\block[]{Conclusion}{\input{conclusion.tex}}
\end{columns}

\block[]{}{\input{reference.tex}}

\end{document}